---
title: "White Wine Quality Data Analysis"
author: "jmonterrubio"
date: "12 de abril de 2016"
output: html_document
---


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(dplyr)
library(psych)
library(scales)
library(memisc)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wines <- read.csv(file = "wineQualityWhites.csv")
# Remove first column (row number)
wines$X <- NULL 
```

# Univariate Plots Section
```{r echo=FALSE, Summary}
dim(wines)
names(wines)
str(wines)
summary(wines)
```

The percentage of alcohol in wine bounds from 8 and 14.20 percent.
About 75% of the wines have a residual sugar value below 10 grams/liter (over 45 are considered sweet).
Some wines have not citric acid.
Mean quality is 5.878, max and min quality are 9 and 3 respectively.


```{r echo=FALSE, Quality}
ggplot(aes(quality), data = wines) +
  geom_histogram(binwidth = 1, colour="black", fill="white") +
  scale_x_continuous(breaks = seq(3, 9, 1))
```

Quality distribution appears unimodal normal distribution with center in 6.

```{r echo=FALSE, Alcohol}
qplot(alcohol, data = wines)
ggplot(aes(alcohol), data = wines) +
  geom_histogram(binwidth = 0.1, colour="black", fill="white") +
  scale_x_continuous(breaks = seq(8, 15, 0.5))

summary(wines$alcohol)
```

Most alcohol values are around 9.5 grades, and follow a right skewed normal distribution.

```{r echo=FALSE, Sulphates}
qplot(sulphates, data = wines)
ggplot(aes(sulphates), data = wines) +
  geom_histogram(binwidth = 0.01, colour="black", fill="white") +
  scale_x_continuous(breaks = seq(0.2, 1.1, 0.1))

summary(wines$sulphates)
```

Most sulphates values are around 0.45 grades, and follow a right skewed normal distribution.

```{r echo=FALSE, pH}
qplot(pH, data = wines)
ggplot(aes(x = pH), data = wines) +
  geom_histogram(binwidth = 0.02, colour="black", fill="white") +
  scale_x_continuous(breaks = seq(2.7, 3.9, 0.1))

summary(wines$pH)
```

pH follow a normal distribution with mean near 3.2

```{r echo=FALSE, Density}
qplot(density, data = wines)
ggplot(aes(x = density), data = wines) +
  geom_histogram(binwidth = 0.0001, colour="black", fill="white") +
  scale_x_log10(breaks = seq(0.985,1.005, 0.002), limits = c(0.985,1.005))

summary(wines$density)
```

Density follow a normal distribution with mean in 0.994 and some outliers over 1.01

```{r echo=FALSE, Total_sulfur_dioxide}
qplot(total.sulfur.dioxide, data = wines)
ggplot(aes(x = total.sulfur.dioxide), data = wines) +
  geom_histogram(binwidth = 5, colour="black", fill="white") +
  scale_x_continuous(breaks = seq(0,275, 25), limits = c(0,275))

summary(wines$total.sulfur.dioxide)
```

The total sulfur dioxide follow a normal distribution with mean in 138.4 and some outliers over 275. The 50% if the data is between 108 and 167

```{r echo=FALSE, Free_sulfur_dioxide}
qplot(free.sulfur.dioxide, data = wines)
ggplot(aes(x = free.sulfur.dioxide), data = wines) +
  geom_histogram(binwidth = 1, colour="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 290, 10), limits = c(0,90))

summary(wines$free.sulfur.dioxide)
```

The free sulfur dioxide follow a normal distribution with mean near 35 and some outliers over 90. The 50% if the data is between 23 and 46

```{r echo=FALSE, Chlorides}
qplot(chlorides, data = wines)
ggplot(aes(x = chlorides), data = wines) +
  geom_histogram(binwidth = 0.02, colour="black", fill="white") +
  scale_x_log10(breaks = c(0.01, 0.015, 0.02, 0.03, 0.05, 0.1, 0.2))

summary(wines$chlorides)

```

Almost 50% of chlorides are between 0.036 and 0.05.

```{r echo=FALSE, Residual_sugar}
qplot(residual.sugar, data = wines)
ggplot(aes(x = residual.sugar), data = wines) +
  geom_histogram(binwidth = 0.02, colour="black", fill="white") +
  scale_x_log10(breaks = c(1, 2, 5, 10, 20))

```

Transformed the long tail data to better understand the distribution of residual sugar. The tranformed residual sugar distribution appears bimodal with the price peaking around 1.5 or so and again at 10 or so. This is one interesting plot.

Based on the context of this feature, it's a good candidate to transform it in a new factorized variable. The description says that it's rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet so our new variable could be:

```{r echo=FALSE, Sugar_taste}
wines <- mutate(wines,
         sugar.taste = ifelse(residual.sugar < 1, 'RARE',
                       ifelse(residual.sugar > 45, 'SWEET', 'NORMAL')))
wines$sugar.taste <- ordered(wines$sugar.taste, levels = c('RARE', 'NORMAL', 'SWEET'))
summary(wines$sugar.taste)
```

Sadly there aren't enought cases in SWEET and RARE to consider this feature.

```{r echo=FALSE, Citric_acid}
qplot(citric.acid, data = wines)
ggplot(aes(x = citric.acid), data = wines) +
  geom_histogram(binwidth = 0.01, colour="black", fill="white") +
  scale_x_continuous(breaks = seq(0, 1.7, 0.1), limits = c(0,0.75))

summary(wines$citric.acid)
```

The citric acid follow a normal distribution with mean near 0.33 and some outliers over 0.9. The 50% if the data is between 0.27 and 0.39.

Setting the binwidth to 0.01 we can see an anormal amount of values around 0.5 (0.49 exactly).

```{r echo=FALSE, Citric_acid_table}
sort(table(wines$citric.acid), decreasing = T)
```

```{r echo=FALSE, Volatile_acidity}
qplot(volatile.acidity, data = wines)
ggplot(aes(x = volatile.acidity), data = wines) +
  geom_histogram(binwidth = 0.01, colour="black", fill="white") +
  scale_x_log10(breaks = seq(0, 1.1, 0.1))
```

Transformed the long tail data to better understand the distribution of volatile acidity. The tranformed volatile acidity distribution appears normal with the acifuty peaking around 0.25

```{r echo=FALSE, Fixed_acidity}
qplot(fixed.acidity, data = wines)
ggplot(aes(x = fixed.acidity), data = wines) +
  geom_histogram(binwidth = 0.1, colour="black", fill="white") +
  scale_x_continuous(breaks = seq(4, 10, 1), limits = c(4,10))

summary(wines$fixed.acidity)
```

The fixed acidity follow a normal distribution with mean near 6.85 and some outliers over 10 and under 4. The 50% if the data is between 6.3 and 7.3

A new feature could be obtained using the acidities. Some references says that the total acidity is the amount of fixed acidity plus the volatile acidity. But the measure of fixed acidity should be setted (for an easier understanding) to just tartaric acid and not all the non-volatile acids so our total acidity is going to be the sum of all the acids.

```{r echo=FALSE, Total_acidity}
wines <- mutate(wines,
         total.acidity = fixed.acidity + volatile.acidity + citric.acid)
qplot(total.acidity, data = wines)
ggplot(aes(x = total.acidity), data = wines) +
  geom_histogram(binwidth = 0.1, colour="black", fill="white") +
  scale_x_continuous(breaks = seq(4, 11, 1), limits = c(4,11))
summary(wines$total.acidity)
```

# Univariate Analysis

### What is the structure of your dataset?

There are 4898 wines in the dataset with 12 features (fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, alcohol and quality).

All the features are numerical, even the quality that is based on a score of 0 to 10. This feature is the easiest one to be factorized for an easy plot interpretations but we are going to mantain both.

```{r echo=FALSE, Factorize_quality}
wines$quality.factor <- ordered(wines$quality, levels = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10))
```

Main thoughts:

Seeing the univariate plots, most of the features follow normal distributions with few variability but some outliers.

### What is/are the main feature(s) of interest in your dataset?

The main feature in the data set is the quality. The main idea is try to predict the quality of a wine. To accomplish this issue let see what is the behaviour of the quality with some other features.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

I've learned, reviewing some internet articles, that a good wine quality is given by this formula: Sweet Taste (sugars + alcohols) <= => Acid Taste (acids) + Bitter Taste (phenols). In the case of white wines, the concentration of phenols (tannins, that gives the red color of the wine) are insignificant. So the interesting features for this analysis will be: fixed.acidity, volatile.acidity, citric.acid, total.acidity, residual.sugar, alcohol and quality.

### Did you create any new variables from existing variables in the dataset?

Total acidity is a combination of all the acids. Also the residual sugar has been categorized but the few quantity of data in some categories made this one useless.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

Residual sugar maybe is the most unusual distribution cause for a better understanding of the data a log10 has been applied and appears a bimodal distribution. The rest of them seems to be normal distributions, some of them right skewed.

# Bivariate Plots Section

```{r echo=FALSE, Correlation_Matrix_All}
cor(wines[, c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 14)])
```

```{r echo=FALSE, Pairs_Panel_All}
set.seed(10)
pairs.panels(wines[, c(4 ,5, 6, 7, 8, 9, 10, 11, 12, 14)], pch=".")
```

Some unexpected correlations appears in the features.  The density seems to be correlated with the residual sugar and with the alcohol. So lets insclude this one in the investigation features.

```{r echo=FALSE, Correlation_Matrix}
cor(wines[, c(4, 8, 11, 12, 14)])
```

```{r echo=FALSE, Pairs_Panel}
set.seed(10)
pairs.panels(wines[, c(4, 8, 11, 12, 14)], pch=".")
```

The main objective is to know how this features affect the wine quality, but first lest see how others features are related.

```{r echo=FALSE, Residual_Sugar_vs_Density}
ggplot(aes(x = residual.sugar, y = density), data = wines) +
  geom_point(color = I('seagreen'),alpha = 0.5, size = 1, position = 'jitter') +
  coord_cartesian(xlim = c(0, quantile(wines$residual.sugar, .95))) +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle('Density by Residual Sugar')
```

Here is a very strong relationship between the residual sugar and the density of the wine. In fact the correlation is 0.84.

```{r echo=FALSE, Density_vs_Alcohol}
ggplot(aes(x = density, y = alcohol), data = wines) +
  geom_point(color = I('seagreen'), alpha = 0.5, size = 1, position = 'jitter') +
  coord_cartesian(xlim = c(quantile(wines$density, .05),
                           quantile(wines$density, .95))) +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle('Alcohol by Density')
```

Also we see a strong relationship between density and alcohol as we could advance with the correlation of -0.78. 

```{r echo=FALSE, Quality_vs_Total_Acidity}
qplot(x = total.acidity, y = quality, data = wines)
ggplot(aes(x = total.acidity, y = quality), data = wines) +
  geom_point(color = I('seagreen'), alpha = 0.5, size = 1, position = 'jitter') +
  coord_cartesian(xlim = c(quantile(wines$total.acidity, .01),
                           quantile(wines$total.acidity, .99))) +
  #geom_smooth(method = "lm", se = FALSE) +
  ggtitle('Quality by Total Acidity')
```

No relationship can be shown, in fact the correlation value for this pairs is -0.13.

```{r echo=FALSE, Total_Acidity_by_Quality}
qplot(total.acidity, data = wines, binwidth = .05, fill = quality.factor) +
  coord_cartesian(xlim = c(quantile(wines$total.acidity, .01),
                           quantile(wines$total.acidity, .99))) +
  guides(fill = guide_legend(reverse = T))
with(wines, by(residual.sugar, quality, summary))
```

```{r echo=FALSE, Quality_vs_Residual_Sugar}
ggplot(aes(x = residual.sugar, y = quality), data = wines) +
  geom_point(color = I('seagreen'), alpha = 0.5, size = 1, position = 'jitter') +
  coord_cartesian(xlim = c(0, quantile(wines$residual.sugar, .95))) +
  #geom_smooth(method = "lm", se = FALSE) +
  ggtitle('Quality by Residual Sugar')
```
```{r echo=FALSE, Residual_sugar_by_Quality}
qplot(residual.sugar, data = wines, binwidth = .1, fill = quality.factor) +
  coord_cartesian(xlim = c(0, quantile(wines$residual.sugar, .95))) +
  guides(fill = guide_legend(reverse = T))
with(wines, by(residual.sugar, quality, summary))
```

Again any relationship between these features as expected.

```{r echo=FALSE, Quality_vs_Alcohol}
ggplot(aes(x = alcohol, y = quality), data = wines) +
  geom_point(color = I('seagreen'), alpha = 0.1, size = 5, position = 'jitter') +
  coord_cartesian(xlim = c(quantile(wines$alcohol, .01),
                           quantile(wines$alcohol, .99))) +
  #geom_smooth(method = "lm", se = FALSE) +
  ggtitle('Quality by Alcohol')
```
```{r echo=FALSE, Alcohol_by_Quality}
qplot(alcohol, data = wines, binwidth = .1, fill = quality.factor) +
  coord_cartesian(xlim = c(quantile(wines$alcohol, .01),
                           quantile(wines$alcohol, .99))) +
  guides(fill = guide_legend(reverse = T))
with(wines, by(alcohol, quality, summary))
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

The main relationships in this bivariate analysis are found related with the alcohol feature. We could see that it has a strong relationship with the density and the residual sugar.

But no single relationship could be found with the quality. Each of the features analyzed aren't somehow related with the quality. This is something we can expected because is not that easy to have a good wine quality, isn't it?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

# Bibliography

## Feature Knowledge

### acidity

http://www.calwineries.com/learn/wine-chemistry/acidity
http://winemakersacademy.com/understanding-wine-acidity/

### volatile acidity

http://extension.psu.edu/food/enology/wine-production/volatile-acidity-in-wine

### citric acid

http://www.calwineries.com/learn/wine-chemistry/wine-acids/citric-acid

### residual sugar

http://www.calwineries.com/learn/wine-chemistry/sugar-in-wine

### alcohol

http://www.calwineries.com/learn/wine-chemistry/alcohol

# Some toughts

The higher the sugar, the higher alcohol
Sweet Taste (sugars + alcohols) <= => Acid Taste (acids) + Bitter Taste (phenols) (tannins just red wines)
Increase fixed acidity decreases the ph
Increase citric acid increases ph